Stateless Model Checking with Data-Race Preemption Points

### 2015-10-20 pldi-16/di.tex
write algorithm 1, GetBestJob

Iterative Deepening depends heavily on state-space estimation \cite{estimation}
to understand which state spaces are likely to complete on time,
in advance of actually testing each interleaving within.
The algorithm's purpose is to make decisions automatically about when to defer exploration of a state space,
so an inexpert user need only provide their total CPU budget as an argument.

@@

For the remainder of the paper, we will abbreviate {\em preemption point} (PP),
{\em model checking} (MC),
and {\em single-state-space MC} (SSS-MC) (i.e., the approach of prior work).

@@

We also identify the {\em minimal} and {\em maximal state space} for each test.
The {\em minimal state space} includes only thread switches arising from normal execution (Figure~\ref{fig:id}).
The {\em maximal state space} is simply the one explored by SSS-MC: all statically-available PPs are enabled.
However, should new data-race PPs be added during a test, the new maximal state space will be the one including those as well.

@@

This algorithm is the heart of Iterative Deepening; we list it in Figure~\ref{fig:shouldworkblock}.
The algorithm takes into account... TODO
Though it has worst-case performance of $O(mn)$ in the number of pending and blocked jobs,
in practice we find the non-constant portion runs very infrequently and is negligible compared to the exponentially-sized state spaces.

>>>>

Iterative Deepening depends heavily on state-space estimation \cite{estimation}
to understand which state spaces are likely to complete on time,
in advance of actually testing each interleaving within.
The purpose is to make decisions automatically about when to defer exploration of a state space,
so an inexpert user need only provide their total CPU budget as an argument,
and to enable completing appropriately-sized state spaces within that budget.

@@

For the remainder of the paper, we will abbreviate {\em preemption point} (PP),
{\em model checking} (MC),
{\em single-state-space model checking} (SSS-MC) (i.e., the approach of prior work),
and {\em state space estimate} (ETA).

@@

We also identify the {\em minimal} and {\em maximal state space} for each test.
The {\em minimal state space} includes only thread switches arising from normal execution (Figure~\ref{fig:id}).
The {\em maximal state space} is simply the one explored by SSS-MC: all statically-available PPs are enabled.
%However, should new data-race PPs be added during a test, the new maximal state space will be the one including those as well.

@@

% TODO: put colour in this code
\newcommand\PendingJobs{\ensuremath{\mathcal{P}}}
\newcommand\SuspendedJobs{\ensuremath{\mathcal{S}}}
\newcommand\GetETA[1]{ETA(#1)}
\newcommand\GetPPSet[1]{PPSet(#1)}
\begin{algorithm}[t]
	\SetKwInOut{Input}{Input}
	%\textbf{Function} GetBestJob($j_0$, PendingJobs, SuspendedJobs): \\
	\Input{$j_0$, the currently-running job}
	%\Input{$eta$, $j_0$'s predicted completion time}
	\Input{\PendingJobs, the list of pending jobs, sorted decreasingly by heuristic priority}
	\Input{\SuspendedJobs, the list of already-suspended jobs, sorted increasingly by ETA}
	\If{\GetETA{$j_0$} $<$ HeuristicETAFactor $\times$ TimeLeft()}{
		// common case: job will probably finish on time \\
		return $j_0$
	}
	\ForEach{job $j_P \in$ \PendingJobs}{
		// don't run a pending job if a subset of it is already suspended; it would find the same fate! \\
		\If {$\forall j_S \in$ \SuspendedJobs, \GetPPSet{$j_S$} $\not\subset$ \GetPPSet{$j_P$}}{
			return $j_P$
		}
	}
	%// no pending jobs; maybe resume a suspended job \\
	\ForEach{job $j_S \in$ \SuspendedJobs}{
		\If{\GetPPSet{$j_0$} $\not\subset$ \GetPPSet{$j_S$}
			$\land$
			\GetETA{$j_0$} $>$ \GetETA{$j_S$}}{
			// $j_S$ is tempting, but if a subset of it is also suspended, don't run the larger one first \\
			\If{$\forall j_{S2} \in$ \SuspendedJobs, \GetPPSet{$j_{S2}$} $\not\subset$ \GetPPSet{$j_S$}}{
				return $j_S$
			}
		}
	}
	// $j_0$'s ETA was bad, but nothing else was better \\
	return $j_0$
	\caption{Suspending exploration of a too-large state space in favour of a potentially smaller one.}
	\label{alg:shouldworkblock}
\end{algorithm}

With a limited CPU budget, we must avoid running tests likely to time out.
When a MC instance reports too high ETA (greater than some heuristic factor times the time remaining),
we compare other yet-untested PP sets to find another job more likely to complete in time.
This algorithm is the heart of Iterative Deepening; we list it in Algorithm~\ref{alg:shouldworkblock}\footnote{
Though its worst-case performance is $O(mn)$ in the number of pending and suspended jobs,
in practice the non-constant portion runs very infrequently
and is negligible compared to the exponentially-sized state spaces.}.
The algorithm takes into account... TODO
